Business Analytics - Finding Groups Within the Data

### 2016-04-05 finding_groups/finding_groups.Rmd
radar plot, notebook format

---
title: "Business Analytics 1 : Finding Groups Within the Data"
author: "Ming-Yu Liu"
date: "November 11, 2015"
output: 
  html_document: 
    number_sections: yes
    theme: united
    toc: yes
---

@@

Let's read in the dataset to get a feel of what it looks like, before going further. Oh, and the link to the datasets used in this documentation is located at the bottom of the report.

```{r, message=FALSE}

# environment setting
library(dplyr)
library(caret)
library(ggplot2)
library(reshape2)
setwd("/Users/ethen/Business-Analytics/1_finding_groups")
rm( list = ls() )

data_sku <- read.csv("data/DATA_2.01_SKU.csv")
head(data_sku)

@@

ggplot( data_sku, aes( CV, ADS ) ) + geom_point() + 
labs( title = "SKU Example", y = "Average Daily Sales", x = "Coefficient of Variation" ) + 
geom_hline( yintercept = 4, color = "blue" ) + 
geom_vline( xintercept = .2, color = "blue" )

@@

# hierarchical clustering, with k (clustering number) = 3 
d <- dist( scale(data_sku), method = "euclidean" )
cluster <- hclust( d, method = "ward.D" )
( data_sku$groups  <- as.factor( cutree( cluster, k = 3 ) ) )

@@

ggplot( data_sku, aes( CV, ADS, color = groups ) ) + geom_point() + 
labs( title = "SKU Example", y = "Average Daily Sales", x = "Coefficient of Variation" )

@@

data_hr <- read.csv("data/DATA_2.02_HR.csv")
str(data_hr)

@@

```{r}
# scale the data (you don't really need to do this when comparing correlations)
# this is used for latter plotting and applying clustering
hr_scaled <- data.frame( scale(data_hr) )

# find variables that have correlations higher than 0.8 and print out variable names
findCorrelation( cor(hr_scaled), cutoff = 0.8, names = TRUE )

@@

```{r, message=FALSE}

ggplot( melt( hr_scaled[ , 2:3 ] ), aes( value, fill = variable ) ) + 
geom_histogram( alpha = .4, position = "identity" )

@@

d <- dist( hr_scaled[ , -1 ], method = "euclidean" ) 
cluster <- hclust( d, method = "ward.D" )
# dendogram
# plot(cluster)

data_hr$groups <- cutree( cluster, k = 4 )

# median aggregation 
hr_agg <- aggregate( .~ groups, data = data_hr, FUN = median )

# count the cluster proportion size 
tl <- table(data_hr$groups)
hr_agg$proportion <- tl / sum(tl)

# order by proportion 
mutate( hr_agg, proportion = tl / sum(tl) ) %>%
arrange( desc(proportion) )

@@

When presenting the upper result table, one visualization you can use is the radar chart. However since the current version of the ggplot2 library still does not support actual radar chart with a simple function call (as least not to my knowledge) and each variable of the data is recorded with different scales. Therefore, the visualization after using the max-min normalization method on the recorded score may look a bit hacky ( the `scale` function produces negative numbers and that does not sound very intuitive for radar charts, thus a different kind of normalization is used here ). As with this normalization, there will be values of 0.

@@

Normalize <- function(x) ( x - min(x) ) / ( max(x) - min(x) )

# exclude the proportion, groups and store the groups
hr_agg_2 <- hr_agg[ , c( -1, -8 ) ]
groups <- hr_agg$groups

# transform to long format for plotting
hr_agg_melt <- apply( hr_agg_2, 2, Normalize ) %>%
			   data.frame() %>%
			   mutate( groups = groups ) %>%
               melt( id.vars = "groups" )

# radar chart
ggplot( hr_agg_melt, aes( x = variable  ) ) + 
geom_bar( aes( y = value, fill = groups ), stat = "identity" ) + 
coord_polar() + 
facet_wrap( ~ groups ) +
ggtitle( "HR" ) + 
theme( legend.position = "none",  
	   axis.text.y = element_blank(), 
	   axis.text.x = element_text( color = "black" ), 
	   axis.ticks  = element_blank(),
	   axis.title  = element_blank() )

@@

**Takeaways:**

@@

This wraps up our example of where clustering algorithms can be used to improve your businesses efficiency. You can find all the code (finding_groups.R) and datasets (in the data folder) used in this documentation [here](https://github.com/ethen8181/Business-Analytics/tree/master/1_finding_groups).

>>>>

---
title: "Business Analytics - Finding Groups Within the Data"
author: "Ming-Yu Liu"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: pygments
---

@@

Let's read in the dataset to get a feel of what it looks like, before going further. Oh, and the link to the datasets used in this documentation is located at the bottom of the documentation.

```{r, message=FALSE, warning=FALSE}
library(caret)
library(scales)
library(ggradar)
library(ggplot2)
library(data.table)

setwd("/Users/ethen/Business-Analytics/finding_groups")
rm( list = ls() )

# ----------------------------------------------------------
#              example 1 : supply chain
# ----------------------------------------------------------
data_sku <- fread("data/DATA_2.01_SKU.csv")
head(data_sku)

@@

ggplot( data_sku, aes( CV, ADS ) ) + geom_point() +  
geom_hline( yintercept = 4, color = "blue" ) + 
geom_vline( xintercept = .2, color = "blue" ) + 
labs( x = "Coefficient of Variation", y = "Average Daily Sales", 
	  title = "SKU Example" )

@@

# hierarchical clustering, with k (clustering number) = 3 
d <- dist( scale(data_sku), method = "euclidean" )
cluster <- hclust( d, method = "ward.D" )
data_sku[ ,  groups := as.factor( cutree( cluster, k = 3 ) ) ]
head(data_sku)

@@

ggplot( data_sku, aes( CV, ADS, color = groups ) ) + geom_point() + 
labs( x = "Coefficient of Variation", y = "Average Daily Sales", 
	  title = "SKU Example" )

@@

data_hr <- fread("data/DATA_2.02_HR.csv")
head(data_hr)

@@

```{r}
# scale the data (you don't really need to do this when comparing correlations)
# this is used for latter plotting and applying clustering
scaled <- scale(data_hr)
data_hr_scaled <- data.table(scaled)

# find variables that have correlations higher than 0.8 and print out variable names
findCorrelation( cor(data_hr_scaled), cutoff = 0.8, names = TRUE )

@@

```{r, message=FALSE, warning=FALSE}

ggplot( melt( data_hr_scaled[ , c( "LPE", "NP" ), with = FALSE ] ), 
		aes( value, fill = variable ) ) + 
geom_histogram( alpha = .4, position = "identity" )

@@

# clustering 
d <- dist( data_hr_scaled[ , -1, with = FALSE ], method = "euclidean" ) 
cluster <- hclust( d, method = "ward.D" )
# plot(cluster) # dendogram
data_hr_scaled[ ,  groups := as.factor( cutree( cluster, k = 4 ) ) ]

# median aggregation 
hr_agg <- data_hr_scaled[ , lapply( .SD, median ), by = groups ]

# order by cluster's proportion size
t <- table(data_hr_scaled$groups)
hr_agg[ , proportion := t / sum(t) ][ order(-proportion) ]

@@

When presenting the upper result table, one visualization you can use is the radar chart.

- To use the `ggradar` function you will need some extra fonts for it to work, see instructions [here](https://github.com/ricardo-bion/ggradar) to get it. Currently, it does not support facetting and the first column has to be the group (future work).
- Each variable of the data is recorded with different scales. And normalizing them via the `scale` function produces negative numbers and that does not sound very intuitive for radar charts, thus a different kind of normalization is used here, namely the max-min normalization. With this method, you minus each column by its min value and divide it by the difference between the maximum and minimum value `( x - min(x) ) / ( max(x) - min(x) )`. Though the recorded score may look a bit odd as with this normalization, there will be values of 0.

@@

hr_agg2 <- hr_agg[ , -c( "groups", "proportion" ), with = FALSE ]
hr_agg2 <- hr_agg2[ , lapply( .SD, rescale ) ]
hr_agg2[ , groups := as.factor( paste0( "group", 1:nrow(hr_agg2) ) ) ]
name <- colnames(hr_agg2)
setcolorder( hr_agg2, c( "groups", name[ name != "groups" ] ) )

# radar plot
ggradar(hr_agg2)

@@

## Takeaways

@@

This wraps up our example of where clustering algorithms can be used to improve your businesses efficiency. You can find all the code (finding_groups.R) and datasets (in the data folder) used in this documentation [here](https://github.com/ethen8181/Business-Analytics/tree/master/finding_groups).
